Best practices for forecasting changes in biodiversity: an example using breeding birds

### 2017-08-18 manuscript/draft.Rmd
Edits to Figure 1 legend

Biodiversity forecasts are important for conservation, management, and
evaluating how well current models characterize natural systems. While the
number of forecasts for biodiversity are increasing, there is little information
available on how well these forecasts work. Most biodiversity forecasts are not
evaluated to determine how well they predict future diversity, fail to account
for uncertainty, and do not use time-series data that captures the actual
dynamics being studied. We addressed these limitations by using best practices
to explore our ability to forecast the species richness of breeding birds in
North America. We used hindcasting to evaluate six different modeling approaches
for predicting richness. Hindcasts for each method were evaluated annually for a
decade at `r numbers$N_predict_sites` sites distributed throughout the
continental United States. While each model could explain most of the variance
in richness, none of them consistently outperformed a baseline model that
predicted constant richness at each site. In particular, we found no evidence
that current methods (such as species distribution models) can successfully turn
spatial data into useful temporal predictions about biodiversity at decadale
time-scales. The best practices implemented in this study directly influence the
forecasts, the relative performance of different modeling approaches, and the
conclusions about the current state of biodiversity forecasting . To facilitate
the rapid improvement of biodiversity forecasts, we emphasize the value of
specific best practices in making forecasts and evaluating forecasting methods.

@@

![Example predictions from our six models for a single site. The data points
from 1982 through 2003, connected by solid lines, were used for training all six
models; the remaining points were used for evaluating the models' forecasts. In
each panel, point estimates for each year are shown with lines; the darker
ribbon indicates 1 standard deviation of uncertainty, and the lighter ribbon
indicates 95% confidence intervals. **A.** The single-site models models were
trained using one site at a time, based solely on observed richness values
shown. The first two models ("average" and "naive") served as our simple
forecasting baselines. **B.** The environmental models were trained to predict
richness based on elevation, climate, and NDVI; unlike the single-site models,
the environmental models' predictions change from year to year as environmental
conditions change.](../figures/model_predictions.png)

>>>>

Biodiversity forecasts are important for conservation, management, and
evaluating how well current models characterize natural systems. While
biodiversity forecasts are common, there is little information available on how
well these forecasts work. Most biodiversity forecasts are not evaluated in
terms of their ability to predict future diversity, fail to account for
uncertainty, and do not use time-series data that captures the actual dynamics
being studied. We compiled a list of ten best practices for forecasting
biodiversity, and used hindcasting to evaluate six different modeling approaches
for predicting breeding bird species richness. Each method was evaluated at `r
numbers$N_predict_sites` sites distributed throughout the continental United
States. Forecasts were evaluated for up to a decade after the final training
observations. While each model could explain most of the variance in richness,
none of them consistently outperformed a baseline model that simply predicted
constant richness at each site. In particular, we found no evidence that current
methods (such as species distribution models) can successfully turn spatial data
into useful temporal predictions about biodiversity. This result may call into
question the validity of many high-profile forecasts from the last 15 years.
Most of the best practices implemented in this study (e.g., evaluating
confidence intervals as well as point estimates) directly influence the outcome
of the forecasts; some even led to large changes in the relative performance of
different modeling approaches. To facilitate the rapid improvement of
biodiversity forecasts, we emphasize the value of specific best practices in
making and evaluating different forecasting methods.

@@

![Example predictions from six forecasting models for a single site. Data from
1982 through 2003, connected by solid lines, were used for training the models;
the remaining points were used for evaluating the models' forecasts. In each
panel, point estimates for each year are shown with lines; the darker ribbon
indicates the 68% prediction interval (1 standard deviation of uncertainty), and
the lighter ribbon indicates the 95% prediction interval. **A.** Single-site
models were trained independently on each site's observed richness values. The
first two models ("average" and "naive") served as baselines. **B.** The
environmental models were trained to predict richness based on elevation,
climate, and NDVI; the environmental models' predictions change from year to
year as environmental conditions change.](../figures/model_predictions.png)
